### 2016-08-28 main_train_stacking.Rmd

---
title: "ecmlpkdd2016"
author: "Andy CHung"
date: "22 May 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Config
```{r results='hide', message=FALSE, warning=FALSE}
rm(list = ls())  # Remove all objects
gc()             # Garbage Collection
options(scipen=99)
setwd("~")

library(xgboost)    # xgboost for Machine Learning
library(data.table) # read file
library(readr)

```

# 2Folds stacking
```{r}
# Number of stack
NUM_STACK = 11

## config
N_FOLD = 5
ETA = 0.3
EARLY_STOP = 25
LAMBDA = 30
SUB_SAMPLE = 0.95
MAX_DEPTH = 3
MODEL_ID = 7

params = list(booster = "gbtree", 
              objective = "binary:logistic",
              max_depth = MAX_DEPTH,
              lambda = LAMBDA,
              eta = ETA,
              subsample = SUB_SAMPLE,
              colsample_bytree = 0.7,
              colsample_bylevel = 0.7,
              seed = 6174)

######### Load Data
user_2014_month_1_6 = read_csv("Documents/github/ecmlpkdd2016_final/features/user_2014_month_1_6_new.csv")
user_2014_month_7_12 = read_csv("Documents/github/ecmlpkdd2016_final/features/user_2014_month_7_12_new.csv")
user_2015 = read_csv("Documents/github/ecmlpkdd2016_final/features/user_2015_preprocess_new.csv")

### stacking vectors
counter = 0
for (i in list.files("Documents/github/ecmlpkdd2016_final/cache/2014-1-6_vector/")){
  if (counter == NUM_STACK){
    break
  }
  print(i)
  tmp_f = read_csv(paste0("Documents/github/ecmlpkdd2016_final/cache/2014-1-6_vector/",i))
  names(tmp_f) = c("USER_ID", i)
  user_2014_month_1_6 = merge(user_2014_month_1_6, tmp_f, by= "USER_ID", all.x=T)
  counter = counter+1
}
counter = 0
for (i in list.files("Documents/github/ecmlpkdd2016_final/cache/2014-7-12_vector/")){
  if (counter == NUM_STACK){
    break
  }  
  print(i)
  tmp_f = read_csv(paste0("Documents/github/ecmlpkdd2016_final/cache/2014-7-12_vector/",i))
  names(tmp_f) = c("USER_ID", i)
  user_2014_month_7_12 = merge(user_2014_month_7_12, tmp_f, by= "USER_ID", all.x = T)
  counter = counter+1
}

user_2014 = rbind(as.matrix(user_2014_month_1_6),
                  as.matrix(user_2014_month_7_12))
user_2014 = data.frame(user_2014_month_1_6)

train = user_2014
train_target = train$TARGET_TASK_2
train$TARGET_TASK_2 = NULL
train$USER_ID = NULL

## Init xgboost data frame
set.seed(6174)
dtrain = xgb.DMatrix(data = data.matrix(train),
                     label = train_target)

## CV for small data set
model_cv = xgb.cv(params, dtrain, nrounds = 50000, nfold=N_FOLD, 
      eval_metric = 'auc', maximize=T,
      early.stop.round = EARLY_STOP, print.every.n = 1)

```

```{r}
```